### 2015-01-13 ebscoCompleteListing.Rmd

---
title: "EbscoHistory"
author: "Brian Perron"
date: "January 13, 2015"
output: html_document
---

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r eval=FALSE}
library(dplyr)
library(stringr)
library(xlsx)
library(gdata)

#ebscoFullListing.df <- read.xls("/Users/beperron/Git/SocialWorkResearch/Data/completeListing.xls", sheet=1, header=TRUE)
#save(ebscoFullListing.df, file="ebscoFullListing.R")
```

```{r}
load("ebscoFullListing.R")
ebscoFullListing.df <- ebscoFullListing.df[ebscoFullListing.df$SourceType == "Academic Journal",]
ebscoFullListing.df$IndexingStart <- format(as.Date(ebscoFullListing.df$IndexingStart, "%m/%d/%Y"), "%Y")
ebscoFullListing.df$IndexingEnd <- format(as.Date(ebscoFullListing.df$IndexingStart, "%m/%d/%Y"), "%Y")
barplot(table(as.factor(ebscoFullListing.df$IndexingStart)))
```

#Clean up dates
ebsco.df$IndexingStart <- as.character(ebsco.df$IndexingStart)

ebsco.df$IndexingStart <- as.Date(ebsco.df$IndexingStart, format="%m/%d/%Y")

ebsco.df$IndexingStop <- as.character(ebsco.df$IndexingStop)

#Drop journals with a date from the 1800's
ebsco.df$IndexingStop <- ifelse(nchar(ebsco.df$IndexingStop) ==10, NA, ebsco.df$IndexingStop)
#ebsco.df$IndexingStop <- ifelse(nchar(ebsco.df$IndexingStop) == 0, NA, ebsco.df$IndexingStop)

x.start <- as.Date(ebsco.df$IndexingStart)
x.stop <- ebsco.df$IndexingStop

x.stop <- ebsco.df[complete.cases(ebsco.df$IndexingStop),]

as.Date(x.stop$IndexingStop)

x.startformat(ebsco.df$IndexingStart, format = "%Y/%m/%d")