### 2015-01-14 Analysis.Rmd
Updated files. All functional. Growth of science compared to social work functional.

---
title: '100 Years of Social Work Research: A Data Science Perspective'
output: html_document
---

# Data collection

+ Article record retrieval date:  January 8, 2015
Database: PsychInfo and Social Science Index (via EbscoHost Platform)

+ Search terms:  SO "social work" or SO "social welfare" or SO "social casework" or "social case work" or SO "social service" or "social services" or SO "human service" or SO "human services" or SO "social development" or SO "social environment"

+ Search limiters:  Publication Type: Peer Reviewed Journal; Document Type: Journal Article

+ Search modes: Boolean/Phrase

+ Search results were exported from EbscoHost Platformon in a _generic bibliographic format_.  This is an unstructured text file (*.txt) that was processed using in R.

+ EbscoHost full journal listing retrieval date:  January 11, 2015 (http://www.ebscohost.com/title-lists)

# Data Pre-Processing
```{r initialization_dataprocessing, warning=FALSE, message=FALSE, eval=FALSE}
# Clear workspace
rm(list=ls())

# Read BWR functions for Mac OS
source("/Users/beperron/Git/BibWrangleR/functions/piWrangleR.R")
source("/Users/beperron/Git/BibWrangleR/functions/packages.R")
# Set the path where original raw data are stored
setwd("/Users/beperron/Git/SocialWorkResearch/Data")

# Set the working directory to store files created by BWR functions
path <- "/Users/beperron/Git/SocialWorkResearch/Data"

# Wrangle the data with the BWR function suite
#piBWR.f(csv=FALSE, path=path)
#save(pi.df, file = "piArticles.R")
```

```{r initialization_analysis, message=FALSE, warning=FALSE, comment=NA, echo=FALSE}
rm(list=ls())
setwd("/Users/beperron/Git/SocialWorkResearch/Data")
source("/Users/beperron/Git/BibWrangleR/functions/ggsurv.R")
load("ebscoFullListing.R")
load("piArticles.R")
load("socialWorkJournals.R")
library(dplyr)
library(ggplot2)
library(gridExtra)
library(survival)
library(grid)
library(png)
library(reshape2)

# Inspect dimensions of the data file (Rows X Columns)
dim(pi.df)

# Inspect variable names of the data file
names(pi.df)

# How many unique article titles?  Ebsco Results of most current search is $n=XXXX$. Do not proceed with analysis if the output does not match this result.  
length(which(pi.df$attributes == "TI"))
```

Journals excluded from search results.  _Social Work_ today excluded because it is of a magazine format.

```{r}
pi.df <- pi.df %>% filter(
        record != "Journal of Applied Social Sciences" &
        record != "Early Child Development and Care" &
        record != "The Clinical Supervisor" &
        record != "Children and Youth Services Review" &
        record != "General Hospital Psychiatry" &
        record != "Canadian Journal on Aging" &
        record != "Canadian Journal of Community Mental Health" &
        record != "Behavior Modification" &
        record != "Employee Assistance Quarterly" &
        record != "Journal of Applied Behavioral Science" & 
        record != "The Scientific Review of Mental Health Practice: Objective Investigations of Controversial and Unorthodox Claims in Clinical Psychology, Psychiatry, and Social Work" &
        record != "Systems Research and Behavioral Science" & 
        record != "Journal of Applied Rehabilitation Counseling" & 
        record != "Journal of Educational & Psychological Consultation" &
        record != "Journal of Community & Applied Social Psychology" &                                          
        record != "Evaluation and Program Planning" &                                                                  
        record != "Journal of Nonverbal Behavior" &
        record != "Brain and Cognition" &
        record != "Early Education and Development" & 
        record != "American Journal of Preventive Medicine" &
        record != "Social Work Today")
```

Two journal titles were merged because one set of records was a logical abbreviation of another set of records.

```{r}

pi.df$record <- ifelse(pi.df$record == "Practice", "Practice: Social Work in Action"
                , pi.df$record)
pi.df$articleID <- ifelse(pi.df$articleID == 74, 42, pi.df$articleID)

pi.df$record <- ifelse(pi.df$record == "Eugenics & Social Welfare Bull.", "Eugenics and Social Welfare Bulletin", pi.df$record)
pi.df$articleID <- ifelse(pi.df$articleID == 28, 32, pi.df$articleID)
```

# Preliminary Analyses

All analyses provided here are not in publishable format, but presented to determine the basis of a storyline.  All tables and figures will be styled according to a target journal.

## Summary of all journal titles included in the analysis

To do:

+ In the table, include the first and last years of indexing, and the numbers of articles.  
+ Write the script to output a *.csv file so that it can be included as an Appendix without any formatting.

```{r unique_titles, echo=FALSE}
unique.titles <- filter(pi.df, attributes == "SO")

# Number of unique titles
number.journals <- length(unique(unique.titles$record))

# Unique titles
journals.unique <- unique(unique.titles$record)
#write.csv(journals.unique, "journals.csv")

number.journals
journals.unique
socialWorkJournals <- journals.unique
#save(journals.unique, file="socialWorkJournals.R")
```

## Count of unique journal titles over time

To do:
+ Integrate the full journal listing of EbscoHost 
+ Plot overall number of journals over time 
+ Plot overall number of journals by major discipline category as defined by EbscoHost classification -- use sociology, anthropology, economics, psychology, and political science

```{r unique_titles_year, message=FALSE, comment=NA, warning=FALSE, echo=FALSE}
journals.year <- tbl_df(pi.df)

year <- journals.year %>%
        filter(attributes == "YR") %>%
        select(id = articleID, year = record)

journals <- journals.year %>%
        filter(attributes == "SO") %>%
        select(id = articleID, journal.title = record)

n.journals.year <- journals %>% 
        left_join(year) %>%
        group_by(year) %>%
        distinct(journal.title) %>%
        summarise(n = n())

journal.count <- ggplot(n.journals.year, aes(as.numeric(year), y=n, group=1)) + 
    geom_line(colour="darkblue") +
    #geom_point(colour="red") + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
    xlab("Publication year") + 
    ylab("Number") + 
    ggtitle("Number of Journals Over Time") + 
    scale_x_continuous(breaks=seq(1914, 2014, 5))

journal.count
```

## Growth of Social Work compared to growth of the social sciences

```{r}
ebscoFullListing.df <- ebscoFullListing.df[ebscoFullListing.df$SourceType == "Academic Journal",]
ebscoFullListing.df$IndexingStart <- format(as.Date(ebscoFullListing.df$IndexingStart, "%m/%d/%Y"), "%Y")
ebscoFullListing.df$IndexingStop <- format(as.Date(ebscoFullListing.df$IndexingStop, "%m/%d/%Y"), "%Y")
ebJournals.df <- ebscoFullListing.df %>% filter(SourceType == "Academic Journal") %>%
    select(Discipline, IndexingStart, IndexingStop, PublicationName, ISSN) %>%
    mutate(IndexingStart = as.numeric(IndexingStart), IndexingStop = as.numeric(IndexingStop)) %>%
    filter(Discipline == "Sociology" | Discipline == "Political Science & Studies"| 
           Discipline == "Psychology"| Discipline == "Economics" | Discipline == "Anthropology")

ebJournals.df.unique <- ebJournals.df %>% select(ISSN)
ebJournals.df.unique <- !duplicated(ebJournals.df.unique)
ebJournals.df <- ebJournals.df[ebJournals.df.unique, ]

ebJournals.df <- ebJournals.df[!(ebJournals.df$PublicationName %in% journals.unique),]

ebJournals.df <- select(ebJournals.df, -PublicationName, -ISSN) %>%
    mutate(journalID = c(1:nrow(ebJournals.df))) %>%
    mutate(journalID = as.factor(journalID)) %>% 
    select(-Discipline)

#ebJournals.melted <- melt(ebJournals.df)

#ebJournals.overtime.start <- ebJournals.melted %>% 
#    filter(value = !is.na(value)) %>%
#    group_by(Discipline, value) %>%
#    summarise(N.start = n())

#ebJournals.overtime.end <- ebJournals.melted %>% 
#    filter(variable == "IndexingStop"  & !is.na(value)) %>%
#    group_by(Discipline, value) %>%
#    summarise(N.end = n())

#ebjournals.joined <- left_join(ebJournals.overtime.start, ebJournals.overtime.end) %>%
#    mutate(N.end = ifelse(is.na(N.end), 0, N.end), Cumulative = cumsum(N.start)) %>%
#    mutate(N.adjusted = Cumulative - N.end)

ebJournals.melted.d <- melt(ebJournals.df)

ebJournals.overtime.start.d <- ebJournals.melted.d %>% 
    filter(value = !is.na(value)) %>%
    group_by(value) %>%
    summarise(N.start = n())

ebJournals.overtime.end.d <- ebJournals.melted.d %>% 
    filter(variable == "IndexingStop"  & !is.na(value)) %>%
    group_by(value) %>%
    summarise(N.end = n())

ebjournals.joined.d <- left_join(ebJournals.overtime.start.d, ebJournals.overtime.end.d) %>%
    mutate(N.end = ifelse(is.na(N.end), 0, N.end), Cumulative = cumsum(N.start)) %>%
    mutate(N.adjusted = Cumulative - N.end)

journal.time.reduced <- n.journals.year %>% filter(year >= 1950)
ggplot() + 
    geom_line(data = ebjournals.joined.d, aes(x=value, y=N.adjusted), group=1, colour="purple") + 
    geom_line(data = journal.time.reduced, aes(as.numeric(year), y=n), group=1, colour="darkblue")
```

## What journals published the most number of articles

```{r journal_output, echo=FALSE}

n.so.yr <- filter(pi.df, attributes == "SO" | attributes == "YR")

n.so <- filter(pi.df, attributes == "SO") %>% mutate(title = record) %>% 
        select(-attributes, -record)

n.yr <- filter(pi.df, attributes == "YR") %>% mutate(year = record ) %>% 
        select(-attributes, -record)

n.so.yr <- left_join(n.so, n.yr) %>%
    group_by(title) %>%
    summarise(first = min(year), last = max(year), n.to.date = n()) %>%
    arrange(desc(n.to.date))

# 10 highest number of publications
head(n.so.yr, 10)
```

## What is the lifespan of journals?

To do:
+ Incorporate all EbscoHost journals and the separate disciplinary journals

```{r journal_lifespan, echo=FALSE}
#10 longest running journals 
longest.running <- n.so.yr %>%
       mutate(last = as.numeric(last), first = as.numeric(first), 
              year.diff = last - first) %>%
       arrange(desc(year.diff)) %>%
       select(title, first, last, year.diff) %>%
       mutate(stop = year.diff, event = ifelse(as.numeric(last) != 2013, 1, 0)) %>%
       select(title, stop, event, as.numeric(first))

survival.journals <- survfit(Surv(longest.running$stop+1, longest.running$event) ~ 1)
median.survival <- data.frame(time = c(24,24), quant = c(.5,0))

head(longest.running)

ggsurv(survival.journals) + 
    geom_line(data = median.survival, aes(time, quant), linetype="longdash") +
    annotate("segment", x = 32, xend = 27, y = .12, yend = .12, size = .25, arrow =arrow()) +
    geom_text(x = 45, y =.12, label = "median survival", size = 4) + 
    geom_text(x = 75, y =.75, label = "+ = Last year of indexing", size = 4) +
    ylim(0,1) +
    xlim(0,100) + 
    xlab("Survival Time (in years)") +
    ylab("Survival Probability") +
    ggtitle("Survival Probability of Social Work Journals")

```

## What is the number of articles published per year

To do:
+ Plot against major disciplinary categories

```{r articles_per_year, echo=FALSE}
n.articles.year <- filter(pi.df, attributes == "YR") 
year.split <- split(n.articles.year, n.articles.year$record)
year.count <- unlist(lapply(year.split, nrow))
year.count <- year.count[order(names(year.count))]
years <- names(year.count)

df <- data.frame(years, year.count)
rownames(df) <- NULL

plot.article.count <- ggplot(df, aes(as.factor(years), 
                    y = year.count, group=1)) + 
    geom_line(colour="black") +
    #geom_point(colour="red") + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
    xlab("year") + 
    ylab("count") + 
    ggtitle("Number of Studies by Year") +
    scale_x_discrete(breaks=c(seq(1914, 2015, 10))) +
    scale_y_continuous(breaks = c(seq(0, 2000, 250)))

df$years <- as.numeric(as.character(df$years))

plot.article.cumulative <- ggplot(df, aes(x = years, y = cumsum(year.count))) + 
    geom_line() + 
    theme(axis.text.x = element_text(angle=45, hjust=1)) +
    scale_x_continuous(breaks=pretty(df$years)) +
    xlab("year") +  
    ylab("count") +
    scale_x_continuous(breaks = c(seq(1914,2014,10))) +
    scale_y_continuous(breaks = c(seq(0, 25000, 2500))) + 
    ggtitle("Cumulative Frequency")

grid.arrange(plot.article.count, plot.article.cumulative, ncol=2)

# Print most recent ten years
tail(df, 10)
```

# What are the most frequent topic areas (by author specified keywords)?

To do:  
+ Fix script that performs matching -- convert all keywords to lower before matching

```{r keywords_overall, comment=NA, warning=FALSE, message=FALSE, echo=FALSE}
kp.df <- filter(pi.df, attributes == "KP")

subject.terms <- stringr::str_split(kp.df$record, pattern = ";")
subject.terms <- tolower(subject.terms)
subject.terms <- unlist(lapply(subject.terms, function(x) gsub(" ", "", x)))
subject.terms.total <- length(unlist(lapply(subject.terms, 
                        function(x) gsub(" ", "", x))))

subject.terms.unique <- length(unique(subject.terms))

subject.terms.l <- list(subject.terms.total = subject.terms.total,
                      subject.terms.unique = subject.terms.unique)

most.frequent <- as.data.frame(table(subject.terms))

most.frequent <- arrange(most.frequent, desc(Freq))

# Print summary statistics
print(subject.terms.l)

# Print 25 most frequent
head(most.frequent, 25)
```

## Most Frequent Author Keywords

_DO NOT INTERPRET RESULTS HERE THIS SECTION IS NOT WORKING PROPERLY_

For this analysis (to-do), select a few keywords that may tell an interesting story of how they change over time -- something that might be revealing of 'trends.'

```{r keywords_over_time, warning=FALSE, message=FALSE, echo=FALSE}
kp.yr <- filter(pi.df, attributes == "YR") %>%
    mutate(year = as.numeric(record)) %>% select(-record, -attributes)

kp <- pi.df %>%
        filter(attributes == "KP") %>%
        select(articleID = articleID, keyword = record) %>%
        mutate(keyword = gsub(" ", "", keyword))

kp.year <- kp %>% 
         left_join(kp.yr) %>%
         group_by(keyword, year) %>%
         summarise(n = n()) %>%
         arrange(keyword, year, n)

select.keywords <- c("women", "leadership")
select.keywords.df <- kp.year %>% filter(tolower(keyword) %in% select.keywords)

library(plyr) # Fix code so ply loads before dplyr
key.plot.df <- ddply(select.keywords.df, .(keyword), transform, cumulative = cumsum(n))
detach(package:plyr)

#Fix graph to sort the keywords / colour
ggplot(key.plot.df, aes(x = year, y = cumulative, colour = keyword, group = keyword)) + 
    geom_line() +
    ggtitle("Cumulative Sum of Articles by Keywords") +
    ylab("Cumulative Sum") + 
    xlab("Year") + 
    theme(legend.title=element_blank())

```

## Location of Studies

To do:
+ Determine best format / presentation of map

```{r location, comment=NA, warning=FALSE, message=FALSE, echo=FALSE}
LO.df <- filter(pi.df, attributes == "LO") %>%
         group_by(record) %>%
         summarise(Country = unique(record),
                   N = n()) %>%
        arrange(desc(N)) %>% 
        select(-record)

locations <- as.data.frame(table(LO.df))
locations <- arrange(locations, desc(Freq))
locations.unique <- length(unique(locations))

#write.csv(locations, "location.csv")

```

## Location of studies over time

To do:
+ Determine best format / presentation of map

```{r location_over_time, warning=FALSE, message=FALSE, echo=FALSE}
library(RColorBrewer)
year <- filter(pi.df, attributes == "YR") %>%
    mutate(year = as.numeric(record)) %>% select(-record, -attributes)

location.temp <- pi.df %>%
        filter(attributes == "LO") %>%
        select(articleID = articleID, country = record) %>%
        mutate(country = gsub(" ", "", country))

location.year <- location.temp %>% 
         left_join(year) %>%
         group_by(country, year) %>%
         summarise(n = n()) %>%
         arrange(country, year, n) %>%
         mutate(cumsum_n = cumsum(n), log_n = log(n+1))

top.locations <- LO.df %>% 
                 select(Country) %>%
                 slice(1:5)

top.locations$Country <- gsub(" ", "", top.locations$Country)

location.year.top <- location.year %>% filter(country %in% top.locations$Country)

ggplot(location.year.top, aes(x = year, y = log_n, colour = country )) + 
    geom_line() +
    ggtitle("Log Number of Articles by Countries With Highest Scholarly Ouput") +
    xlab("Publication Year") + 
    ylab("log N+1") +
    theme(legend.position=c(.210, .7))

top.locations.no.us <- LO.df %>% 
                select(Country) %>%
                slice(2:10)

top.locations.no.us$Country <- gsub(" ", "", top.locations.no.us$Country)

location.year.top.no.us <- location.year %>% filter(country %in% top.locations.no.us$Country)

ggplot(location.year.top.no.us, aes(x = year, y = cumsum_n, colour = country )) + geom_line() +
    ggtitle("Cumulative Number of Article by Countries with Highest Scholarly Output \n(Excluding the United States)") +
    theme(legend.position=c(.210, .6))
```
## Countries represents in the social work research

To do:
+ Obtain definition of `LO` tag.  
+ Determine how many articles had LO tag (note the issue if not included from Social Science Index)

```{r fig.width=8, echo=FALSE}
Map100YearWithUS <- readPNG("/Users/beperron/Git/SocialWorkResearch/images/Map100Years.png")
grid.raster(Map100YearWithUS)

Map100YearWithoutUS <- readPNG("/Users/beperron/Git/SocialWorkResearch/images/Map100YearsWithoutUS.png")
grid.raster(Map100YearWithUS)
```

## Countries not represented in social work research

```{r fig.width=8, echo=FALSE}
library(png)
library(grid)
img <- readPNG("/Users/beperron/Git/SocialWorkResearch/images/NotRepresented100Years.png")
grid.raster(img)

```

## Number of authors
```{r comment=NA, message=FALSE, warning=FALSE, echo=FALSE}
n.authors.article <- pi.df %>% 
    filter(attributes == "AU") %>%
    select(id = articleID, author= record) %>%
    mutate(id = as.numeric(id))

n_authors.1 <- n.authors.article %>% 
        group_by(id) %>%
        summarise(n = n())

n_authors.1t <- n_authors.1 %>% mutate(n = ifelse(n >=6, 6, n)) %>% 
    group_by(n) %>%
    summarise(Articles = n())

ggplot(n_authors.1t, aes(x = n, y = Articles)) +
    geom_bar(stat="identity", colour="black", fill = "white") +
    geom_text(aes(label = Articles), vjust=-.3, size = 4) +
    scale_x_discrete(limits = c(as.character(1:5), "6+")) + 
    ggtitle("Number of Articles by Author Count") + 
    xlab("Author Count") + 
    ylab("Articles")
```

## Number of authors over time

This figure shows the average number of authors, along with the standard deviation as the ribbon around the average.  Note that there is a possible problem in these data, with a single article listing a huge number.  That can be corrected at a later time.

```{r message=FALSE, comment=NA, warning=FALSE, fig.height=2, fig.height=4, echo=FALSE}
df.2 <- tbl_df(pi.df)
year <- df.2 %>%
        filter(attributes == "YR") %>%
        select(id = articleID, year = record)

authors <- df.2 %>%
        filter(attributes == "AU") %>%
        select(id = articleID, author = record)

n_authors <- authors %>%
        group_by(id) %>%
        summarise(n=n())

n_authors <- n_authors %>% 
        left_join(year) %>%
        group_by(year) %>%
        summarise(median.n = median(n),
                  average.n = mean(n),
                  min.n = min(n),
                  max.n = max(n),
                  std.dev  = sd(n) )

plot.author.count2 <- ggplot(n_authors, aes(as.numeric(year), y=average.n, group=1)) +  
    geom_ribbon(aes(ymin = average.n-std.dev, ymax=average.n+std.dev), colour="lightblue", fill="lightblue", alpha=.40) +
    geom_line(colour="black") + 
    geom_segment(aes(x= 1940, y = 5.5, xend = 1935, yend = 5.5), 
                 arrow=arrow(length=unit(0.25, "cm"))) +
    annotate("text", x = 1951, y = 5.5, label = "1 outlier case of \n24 listed authors", size=3.5) +
    ylab("Average number of authors (black) \nStandard deviation (light blue)") +
    xlab("Publication year") +
    ggtitle("Average Number of Authors per Article Over Time")

plot.author.count2

```

```