### 2015-02-25 archive/ebscoCompleteListing.Rmd
File management.

---
title: "EbscoHistory"
author: "Brian Perron"
date: "January 13, 2015"
output: html_document
---

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r eval=FALSE}
library(dplyr)
library(stringr)
library(xlsx)
library(gdata)

#ebscoFullListing.df <- read.xls("/Users/beperron/Git/SocialWorkResearch/Data/completeListing.xls", sheet=1, header=TRUE)
#save(ebscoFullListing.df, file="ebscoFullListing.R")
```

```{r}
load("/Users/beperron/Git/SocialWorkResearch/Data/ebscoFullListing.R")
ebscoFullListing.df <- ebscoFullListing.df[ebscoFullListing.df$SourceType == "Academic Journal",]
ebscoFullListing.df$IndexingStart <- format(as.Date(ebscoFullListing.df$IndexingStart, "%m/%d/%Y"), "%Y")
ebscoFullListing.df$IndexingStop <- format(as.Date(ebscoFullListing.df$IndexingStop, "%m/%d/%Y"), "%Y")
ebscoFullListing.df$journalID <- c(1:nrow(ebscoFullListing.df))
barplot(table(as.factor(ebscoFullListing.df$IndexingStart)))
```

```{r}
library(reshape2)
library(dplyr)
ebJournals.df <- ebscoFullListing.df %>% filter(SourceType == "Academic Journal") %>%
    select(journalID, Discipline, IndexingStart, IndexingStop, PublicationName, ISSN) %>%
    mutate(journalID = as.factor(journalID), IndexingStart = as.numeric(IndexingStart), IndexingStop = as.numeric(IndexingStop)) %>%
    filter(Discipline == "Sociology" | Discipline == "Political Science & Studies"| 
           Discipline == "Psychology"| Discipline == "Economics" | Discipline == "Anthropology")

unique.journal.titles <- unique(ebJournals.df$ISSN)
ebJournals.df <-ebJournals.df[unique.journal.titles,]

load("socialWorkJournals.R")
cross.check.journals <- ebJournals.df[ebJournals.df$PublicationName %in% socialWorkJournals,]
cross.check.journals

ebJournals.df <- select(ebJournals.df, -PublicationName, -ISSN)

ebJournals.melted <- melt(ebJournals.df)

ebJournals.overtime.start <- ebJournals.melted %>% 
    filter(value = !is.na(value)) %>%
    group_by(Discipline, value) %>%
    summarise(N.start = n())

ebJournals.overtime.end <- ebJournals.melted %>% 
    filter(variable == "IndexingStop"  & !is.na(value)) %>%
    group_by(Discipline, value) %>%
    summarise(N.end = n())

ebjournals.joined <- left_join(ebJournals.overtime.start, ebJournals.overtime.end) %>%
    mutate(N.end = ifelse(is.na(N.end), 0, N.end), Cumulative = cumsum(N.start)) %>%
    mutate(N.adjusted = Cumulative - N.end)

ebJournals.melted.d <- melt(ebJournals.df)

ebJournals.overtime.start.d <- ebJournals.melted.d %>% 
    filter(value = !is.na(value)) %>%
    group_by(value) %>%
    summarise(N.start = n())

ebJournals.overtime.end.d <- ebJournals.melted.d %>% 
    filter(variable == "IndexingStop"  & !is.na(value)) %>%
    group_by(value) %>%
    summarise(N.end = n())

ebjournals.joined.d <- left_join(ebJournals.overtime.start.d, ebJournals.overtime.end.d) %>%
    mutate(N.end = ifelse(is.na(N.end), 0, N.end), Cumulative = cumsum(N.start)) %>%
    mutate(N.adjusted = Cumulative - N.end)

journal.time.reduced <- n.journals.year %>% filter(year >= 1950)
ggplot() + 
    geom_line(data = ebjournals.joined.d, aes(x=value, y=N.adjusted), group=1, colour="purple") + 
    geom_line(data = journal.time.reduced, aes(as.numeric(year), y=n), group=1, colour="darkblue")

#From main analyses
journal.count <- ggplot(n.journals.year, aes(as.numeric(year), y=n, group=1)) + 
    geom_line(colour="black") +
    #geom_point(colour="red") + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
    xlab("Publication year") + 
    ylab("Number") + 
    ggtitle("Number of Journals Over Time") + 
    scale_x_continuous(breaks=seq(1914, 2014, 5))

```