### 2015-09-29 manuscript.Rmd

---
title: 'Examining a Quarter Century of Publishing Trends in Social Work Research:
  A Data Science Perspective'
author: ''
date: "May 26, 2015"
output: html_document
---

`r knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, comment=NA)`

```{r Initialization}
rm(list=ls())
library(dplyr)
library(ggplot2)
library(gridExtra)
source("./Functions/pretty.R")

# Load data files
load("./Data/HistoricalDatabase.RData")
hodge.df <- readLines("./Data/hodge.txt", n = -1)
scopus.df <- read.csv("./Data/scopus.csv")
abb.df <- read.csv("./Data/master.csv")
sangam.df <- read.csv("./Data/sangam.csv")

# Create colour palette for consistent plotting
# http://colorbrewer2.org/?type=diverging&scheme=PuOr&n=4
cp <- c("#fdb864", "#e66101", "#b2abd2", "#5e3c99" )
```

```{r summarySetUp, eval=TRUE}

#=======================================
# Descriptives
#=======================================

numberArticles <- length(which(full.df$attributes == "article"))

numberJournals <- filter(full.df, attributes == "journal") %>%
  summarise(Unique = n_distinct(record))

unique.swHistory <- filter(full.df, attributes == "journal") %>%
    distinct(record)

hodge.swHistory.intersect <- intersect(hodge.df, unique.swHistory$record)

hodge.swHistory.diff <- setdiff(hodge.df, unique.swHistory$record)

```

```{r Table 1 Journal Summary, eval=TRUE }
#=======================================
# Table 1. Journal summary 
#=======================================
# Integrate min - max years for each journal
unique.titles <- filter(full.df, attributes == "journal") %>%
  group_by(record) %>%
  mutate(Hodge = ifelse(record %in% hodge.df == TRUE, "Y", "N")) %>%
  summarise(N = n(), Hodge.list = max(Hodge))

unique.year <- filter(full.df, attributes == "pubYear") %>%
  group_by(record)

n.so <- filter(full.df, attributes == "journal") %>% mutate(title = record) %>% 
    select(-attributes, -record)

n.yr <- filter(full.df, attributes == "pubYear") %>% mutate(year = record ) %>% 
    select(-attributes, -record)

n.so.yr <- left_join(n.so, n.yr) %>%
    group_by(title) %>%
    summarise(first = min(year), last = max(year), n = n()) %>%
    arrange(desc(n)) %>%
    rename(Journal = title)

table1 <- left_join(n.so.yr, abb.df) %>%
  select(-X) %>%
  select(Journal, Abbrev, n, first, last)

sum_Articles <- sum(table1$n)
median_Articles <- median(table1$n)
mean_Articles <- mean(table1$n)
sd_Articles <- sd(table1$n)

```

```{r Figure 1 Small Multiples Paper}
scopus.df < scopus.df %>% filter(Year <= 2013)

# Create an initial data frame from HD
sm.df <- left_join(n.so, n.yr) %>%
    group_by(title, year) %>%
    summarize(N = n()) %>%
    mutate(Overall.N = sum(N), year = as.numeric(year)) %>%
    ungroup() %>%
    arrange(desc(Overall.N), title) %>%
    rename(Year = year, Journal= title)

sm.df.reduced <- sm.df %>% select(-Overall.N)

scopus.df$DataBase <- rep("Scopus", nrow(scopus.df))
DataBase <- rep("HD", nrow(sm.df.reduced))
sm.df.reduced$DataBase <- DataBase
sm.df.reduced <- sm.df.reduced %>% rename(Documents = N) 
sm.df.reduced$Documents <- ifelse(sm.df.reduced$Documents == 0, NA, sm.df.reduced$Documents)  
sm.df.reduced <- na.omit(sm.df.reduced)

combined <- rbind(scopus.df, sm.df.reduced)
combined_reduced <- combined[combined$Journal %in% c("Families in Society", "Social Work", "British Journal of Social Work", "Journal of Gerontological Social Work", "Research on Social Work Practice", "Social Work in Health Care", "International Social Work", "Health & Social Work", "Journal of Social Work Education"), ]

combined_reduced$Journal <- factor(combined_reduced$Journal, levels = c("Families in Society", "Social Work", "British Journal of Social Work", "Journal of Gerontological Social Work", "Research on Social Work Practice", "Social Work in Health Care", "International Social Work", "Health & Social Work", "Journal of Social Work Education") )

# Create Scopus labels for the facets

scopus.df$Year <- as.numeric(scopus.df$Year)
scopus.df <- filter(scopus.df, Year <= 2013)
scopus.df$Year <- as.numeric(scopus.df$Year)
scopus_annotate <- filter(scopus.df, Year <= 2013) %>%
  group_by(Journal) %>%
  summarize(N = pretty(sum(Documents)))

scopus_annotate_paper <-scopus_annotate[scopus_annotate$Journal %in% combined_reduced$Journal, ]
scopus_annotate_paper$label <- paste0("N=", pretty(scopus_annotate_paper$N))

# Create HD Labels for the facets
hd_annotate <- combined_reduced %>% 
  filter(DataBase == "HD") %>%
  group_by(Journal) %>%
  summarize(N = sum(Documents))%>%
  mutate(label = paste0("N=", pretty(N)))

ggplot(combined_reduced,  aes(Year, Documents, colour = DataBase)) +
     geom_line(size = 1, alpha = .80) + 
     facet_wrap(~ Journal) +
  theme_bw() + 
   theme(strip.text.x = element_text(size = 12)) +
  theme(legend.position = "bottom", legend.text = element_text(size = 12)) + 
     scale_colour_discrete(guide = guide_legend(title = "Database"), breaks = c("HD", "Scopus"), labels = c("Historical Database", "Scopus")) + 
  scale_colour_manual(values = c(cp[2], cp[4]), labels = c("Historical Database", "Scopus"), guide = guide_legend(title = "")) + 
  ylab("Number of Articles") + 
  xlab("\nYear") +
     annotate("segment", x = 1992, xend = 1995, y = 140, yend = 140, colour = cp[2], size = 1, alpha = .80) + 
     annotate("segment", x = 1992, xend = 1995, y = 120, yend = 120, colour = cp[4], size = 1, alpha = .80) +
   geom_text(data = scopus_annotate_paper, aes(x = 1999, y =  120, label = label), colour = "black", size = 4) +
  geom_text(data = hd_annotate, aes(x = 1999, y =  140, label = label), colour = "black", size = 4)

```

```{r Figure 1 Small Multiples Supplement}
# Create an initial data frame from HD
sm.df <- left_join(n.so, n.yr) %>%
    group_by(title, year) %>%
    summarize(N = n()) %>%
    mutate(Overall.N = sum(N), year = as.numeric(year)) %>%
    ungroup() %>%
    arrange(desc(Overall.N), title) %>%
    rename(Year = year, Journal= title)

scopus.df$Year <- as.numeric(scopus.df$Year)
scopus.df <- filter(scopus.df, Year <= 2013)

# Join HD with Scopus and join abbreviation names
sm.scopus.merged <-full_join(sm.df, scopus.df, by = c("Journal", "Year")) %>%
    select(Year, Journal, N, Overall.N, Documents)

 sm.scopus.merged.table <- group_by(sm.scopus.merged, Journal) %>%
     summarize(DataBase.N = sum(N, na.rm=TRUE), Scopus.N = sum(Documents, na.rm=TRUE)) %>%
     arrange(desc(DataBase.N))

 abb <- read.csv("/Users/beperron/Git/SocialWorkResearch/Data/master.csv")
 abb$ID <- 1:nrow(abb)
 sm.scopus.merged <- full_join(sm.scopus.merged, abb, by = "Journal") %>% 
     select(-X) %>%
     mutate(Documents = ifelse(Documents == 0, NA, Documents))

 # Re-order the factor by Overall.N
 title <- sm.scopus.merged %>% arrange(desc(Overall.N)) %>% 
     distinct(Abbrev) %>% filter(!is.na(Abbrev)) %>% distinct()

 title.order = title$Abbrev

 sm.scopus.merged$Abbrev <- factor(sm.scopus.merged$Abbrev, levels = (title.order))
 sm.scopus.merged <- group_by(sm.scopus.merged, Journal) %>%
     mutate(Overall.Scopus = sum(Documents, na.rm=TRUE)) %>%
     ungroup()
# 
 # Create labels for the facets
 Overall.N <- distinct(sm.scopus.merged, Abbrev) 
 lab1 <- gsub(" ", "", paste("N=", pretty(Overall.N$Overall.N), sep=""))
 lab2 <- gsub(" ", "", paste("N=", pretty(Overall.N$Overall.Scopus), sep=""))
 lab <- paste(lab1, lab2, sep="\n")
 xpos = rep(1999, 80)
 ypos = rep(200, 80)
 ldata <- data.frame(xpos, ypos, lab, Overall.N$Abbrev) %>% 
     rename(Abbrev = Overall.N.Abbrev)

 ldata <- na.omit(ldata)

 titles.reduced <- unique(sm.scopus.merged$Journal)
 titles.reduced.1 <- titles.reduced[1:40]
 titles.reduced.2 <- titles.reduced[41:length(titles.reduced)] 
 cp <- c("#fdb864", "#e66101", "#b2abd2", "#5e3c99" )

 library(grid)
 library(gridExtra)
# Plotting function    
# 
 plotSmallMultiples.1 <-  ggplot(sm.scopus.merged[sm.scopus.merged$Journal %in% titles.reduced.1, ],  aes(Year, N)) +
     geom_line(colour="#e66101") +
     geom_line(data = sm.scopus.merged[sm.scopus.merged$Journal %in% titles.reduced.1, ], aes(Year, Documents), colour = "#533c99", alpha = .80) + 
     facet_wrap(~ Abbrev) + 
     theme_bw() +
     theme(strip.text.x = element_text(size = 8)) + 
     theme(legend.position = "bottom", 
     panel.grid.minor = element_blank(), 
     axis.text.x = element_text(angle=45, hjust=1, size =8)) + 
     geom_text(data = ldata[1:40, ], aes(x =xpos, y = ypos, label = lab), size =2.5) +
     ylim(c(0,225)) +
     scale_x_continuous(breaks=seq(1989, 2015, 8)) +
     annotate("segment", x = 1989, xend = 1992, y = 215, yend = 215, colour = "#e66101") + 
     annotate("segment", x = 1989, xend = 1992, y = 190, yend = 190, colour = "#533c99") +
     ggtitle("Figure 1. Number of Article Records by Journal (1989 - 2013)")

plotSmallMultiples.2 <- ggplot(sm.scopus.merged[sm.scopus.merged$Journal %in% titles.reduced.2, ],  
     aes(Year, N)) +
     geom_line(colour="#e66101") +
     geom_line(data = sm.scopus.merged[sm.scopus.merged$Journal %in% titles.reduced.2, ], 
     aes(Year, Documents), colour = "#533c99", alpha = .80) + 
     facet_wrap(~ Abbrev) + 
     theme_bw() +
     theme(strip.text.x = element_text(size = 8)) +  
     theme(legend.position = "bottom", 
     panel.grid.minor = element_blank(), 
     axis.text.x = element_text(angle=45, hjust=1, size =8)) + 
     geom_text(data = ldata[41:79, ], aes(x =xpos, y = ypos, label = lab), size =2.5) +
     ylim(c(0,225)) +
     scale_x_continuous(breaks=seq(1989, 2015, 8)) +
     annotate("segment", x = 1989, xend = 1992, y = 215, yend = 215, colour = "#e66101") + 
     annotate("segment", x = 1989, xend = 1992, y = 190, yend = 190, colour = "#533c99") +
     ggtitle("Figure 1 (cont.). Number of Article Records by Journal (1989 - 2013)")
```

```{r Figure 2 Journal Growth, eval = FALSE}
#=======================================
# Journals
#=======================================

n.journals.year <- filter(full.df, attributes == "journal") %>%
    select(articleID, record)

n.year <- filter(full.df, attributes == "pubYear") %>%
    select(articleID, record)

n.journals.year <- left_join(n.journals.year, n.year, by = "articleID", copy=TRUE) %>%
    group_by(record.y) %>%
    summarise(Journal.count = n_distinct(record.x)) %>%
    mutate(record.y = as.numeric(record.y))

exp.047 <- c(29, rep(NA, 24)) 
for(i in 2:25){
    exp.047[i] <- exp.047[i-1] + exp.047[i-1]*.047
}

exp.0321 <- c(29, rep(NA, 24)) 
for(i in 2:25){
    exp.0321[i] <- exp.0321[i-1] + exp.0321[i-1]*.0321
}

n.journals.year$Price <- exp.047
n.journals.year$SWHD <- exp.0321
n.journals.year <- mutate(n.journals.year, Journal.count = as.numeric(Journal.count))
n.journals.year <- as.data.frame(n.journals.year)

years <- c(1989, 2005, 2013, 2013)
points <- c(29, n.journals.year$Journal.count[c(17,25)], max(n.journals.year$exp.05))
annotate.df <- data.frame(years, points)

cols = c("Observed Growth" = cp[1], "3.21% CAGR" = cp[4], "4.7% CAGR" = cp[3])

(figure2 <- ggplot(n.journals.year, aes(record.y, 
    y = Journal.count)) + 
    theme_bw() + 
    theme(axis.text.x = element_text(size=8), 
    title = element_text(size = 10), plot.title=element_text(size=10)) + 
    xlab("Year") + 
    ylab("Journals") + 
    scale_y_continuous(breaks = c(seq(0, 100, 10)), limits = c(20, 95)) +
    theme(axis.title.x = element_text(vjust=-0.5), axis.title.y = element_text(vjust=.75)) + 
    geom_line(aes(x = record.y, y = Price, colour = "Estimated 5% CAGR"),
        data = n.journals.year,   colour = cp[2], alpha = .80) +
    geom_line(aes(x = record.y, y = SWHD, colour = "Estimated 4% CAGR"), 
        data = n.journals.year,  colour = cp[2], alpha = .80) +
    geom_line(aes(colour = "Observed Growth"), colour = cp[4], size = 1.5) + 
    stat_smooth(method = "lm", se = FALSE, colour = cp[3]) +
    annotate("text", x = 2016.5, y = 60, label = "3.2% CAGR", size = 5) +
    annotate("text", x = 2016.5, y = 73, label = "Linear growth", size = 5) + 
    annotate("text", x = 2016.5, y = 90, label = "4.7% CAGR", size = 5) + 
    annotate("text", x = 2017.5, y = 65, label = "Observed growth", size = 5) +
    scale_x_continuous(limits = c(1989, 2020), breaks = c(1989, 1993, 1997, 2001, 2005, 2009, 2013)) +
    #ggtitle("Figure 2. Observed Versus Estimated Growth of Journals (1989-2013)") + 
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
          panel.border = element_blank(), axis.line = element_line(), 
          axis.text.x = element_text(size = 16), 
          axis.text.y = element_text(size = 16), 
          axis.title.x = element_text(size = 16), 
          axis.title.y = element_text(size = 16)) +
    theme(plot.title = element_text(size = 19)))
```

```{r Figure 3 Annual Change, eval=FALSE}
curr <- n.journals.year$Journal.count[-1]
prev <- n.journals.year$Journal.count[1:(length(n.journals.year$Journal.count)-1)]
annualChange.numeric <- c(0, (100 * (curr-prev)/ prev))
annualChange.dichotomous <- ifelse(annualChange.numeric < 0, "Decrease", "Increase")
annualChange.df <- data.frame(annualChange.numeric, annualChange.dichotomous)
annualChange.df$year <- c(1989:2013)

figure3 <- ggplot(annualChange.df, aes(year, annualChange.numeric, fill =annualChange.dichotomous)) + 
    theme_bw() + 
    geom_bar(stat="identity") +
    scale_fill_manual(values = cp[c(3,1)], limits = c("Increase", "Decrease")) + 
    theme(legend.position = c(1,1), legend.justification = c(1,1), legend.title = element_blank()) + 
    ylab("Percent change") +
    xlab("Year")  +
    scale_x_continuous(breaks=seq(1989, 2015, 4)) + 
    scale_y_continuous(breaks = seq(-10, 15, 5)) + 
    geom_hline(yintercept=seq(-5, 15, 5), col="white", lwd=0.4) +
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
          panel.border = element_blank(), axis.line = element_line(), 
          axis.text.x = element_text(size = 16), 
          axis.text.y = element_text(size = 16), 
          axis.title.x = element_text(size = 16), 
          axis.title.y = element_text(size = 16)) + 
    geom_hline(yintercept=0, col="gray", lwd=0.6, alpha =.8) +
    #ggtitle("Figure 3. Annual Percentage Change in the Number Journals (1989 - 2013)") +
    theme(plot.title = element_text(size = 19))

#ggsave("figure3.png", figure3, dpi=250)
```

```{r plot02_SmallMultiples, eval=FALSE}
# Remove 2014 & 2015 
n.so <- filter(full.df, attributes == "journal") %>% mutate(title = record) %>% 
  select(-attributes, -record)

n.yr <- filter(full.df, attributes == "pubYear") %>% mutate(year = record ) %>% 
  select(-attributes, -record)

# Create an initial data frame from HD
sm.df <- left_join(n.so, n.yr) %>%
    group_by(title, year) %>%
    summarize(N = n()) %>%
    mutate(Overall.N = sum(N), year = as.numeric(year)) %>%
    ungroup() %>%
    arrange(desc(Overall.N), title) %>%
    rename(Year = year, Journal= title)

# Read Scopus database
#scopus.df2 <- read.csv("/Users/beperron/Git/SocialWorkResearch/Data/scopus_OldHar#vest.csv")

scopus.df$Year <- as.numeric(scopus.df$Year)
scopus.df <- filter(scopus.df, Year <= 2013)

# Join HD with Scopus and join abbreviation names
sm.scopus.merged <-full_join(sm.df, scopus.df, by = c("Journal", "Year")) %>%
    select(Year, Journal, N, Overall.N, Documents)

 sm.scopus.merged.table <- group_by(sm.scopus.merged, Journal) %>%
     summarize(DataBase.N = sum(N, na.rm=TRUE), Scopus.N = sum(Documents, na.rm=TRUE)) %>%
     arrange(desc(DataBase.N))

 abb.df$ID <- 1:nrow(abb.df)
 sm.scopus.merged <- full_join(sm.scopus.merged, abb.df, by = "Journal") %>% 
     select(-X) %>%
     mutate(Documents = ifelse(Documents == 0, NA, Documents))

 # Re-order the factor by Overall.N
 title <- sm.scopus.merged %>% arrange(desc(Overall.N)) %>% 
     distinct(Abbrev) %>% filter(!is.na(Abbrev)) %>% distinct()

 title.order = title$Abbrev

 sm.scopus.merged$Abbrev <- factor(sm.scopus.merged$Abbrev, levels = (title.order))
 sm.scopus.merged <- group_by(sm.scopus.merged, Journal) %>%
     mutate(Overall.Scopus = sum(Documents, na.rm=TRUE)) %>%
     ungroup()
# 
 # Create labels for the facets
 Overall.N <- distinct(sm.scopus.merged, Abbrev) 
 lab1 <- gsub(" ", "", paste("N=", pretty(Overall.N$Overall.N), sep=""))
 lab2 <- gsub(" ", "", paste("N=", pretty(Overall.N$Overall.Scopus), sep=""))
 lab <- paste(lab1, lab2, sep="\n")
 xpos = rep(1999, 80)
 ypos = rep(200, 80)

 Overall.N <- distinct(sm.df, Journal)

 ldata <- data.frame(xpos, ypos, lab, Overall.N$Abbrev) %>% 
     rename(Abbrev = Overall.N.Abbrev)

 ldata <- na.omit(ldata)

 titles.reduced <- unique(sm.scopus.merged$Journal)
 titles.reduced.1 <- titles.reduced[1:40]
 titles.reduced.2 <- titles.reduced[41:length(titles.reduced)] 
 cp <- c("#fdb864", "#e66101", "#b2abd2", "#5e3c99" )

 library(grid)
 library(gridExtra)
# Plotting function    
# 
 plotSmallMultiples.1 <-  ggplot(sm.scopus.merged[sm.scopus.merged$Journal %in% titles.reduced.1, ],  aes(Year, N)) +
     geom_line(colour="#e66101") +
     geom_line(data = sm.scopus.merged[sm.scopus.merged$Journal %in% titles.reduced.1, ], aes(Year, Documents), colour = "#533c99", alpha = .80) + 
     facet_wrap(~ Abbrev) + 
     theme_bw() +
     theme(strip.text.x = element_text(size = 8)) + 
     theme(legend.position = "bottom", 
     panel.grid.minor = element_blank(), 
     axis.text.x = element_text(angle=45, hjust=1, size =8)) + 
     geom_text(data = ldata[1:40, ], aes(x =xpos, y = ypos, label = lab), size =2.5) +
     ylim(c(0,225)) +
     scale_x_continuous(breaks=seq(1989, 2015, 8)) +
     annotate("segment", x = 1989, xend = 1992, y = 215, yend = 215, colour = "#e66101") + 
     annotate("segment", x = 1989, xend = 1992, y = 190, yend = 190, colour = "#533c99") +
     ggtitle("Figure 1. Number of Article Records by Journal (1989 - 2013)")

plotSmallMultiples.2 <- ggplot(sm.scopus.merged[sm.scopus.merged$Journal %in% titles.reduced.2, ],  
     aes(Year, N)) +
     geom_line(colour="#e66101") +
     geom_line(data = sm.scopus.merged[sm.scopus.merged$Journal %in% titles.reduced.2, ], 
     aes(Year, Documents), colour = "#533c99", alpha = .80) + 
     facet_wrap(~ Abbrev) + 
     theme_bw() +
     theme(strip.text.x = element_text(size = 8)) +  
     theme(legend.position = "bottom", 
     panel.grid.minor = element_blank(), 
     axis.text.x = element_text(angle=45, hjust=1, size =8)) + 
     geom_text(data = ldata[41:79, ], aes(x =xpos, y = ypos, label = lab), size =2.5) +
     ylim(c(0,225)) +
     scale_x_continuous(breaks=seq(1989, 2015, 8)) +
     annotate("segment", x = 1989, xend = 1992, y = 215, yend = 215, colour = "#e66101") + 
     annotate("segment", x = 1989, xend = 1992, y = 190, yend = 190, colour = "#533c99") +
     ggtitle("Figure 1 (cont.). Number of Article Records by Journal (1989 - 2013)")
```

```{r Figure 4 CAGR Comparison, eval=FALSE}

CAGR <- function(start, finish, periods){
  # Function to compute the CAGR given simple returns
  #
  # args:
  #  start = starting value
  #  finish = ending cumulative value
  #  periods = number of time periods over which value was accumulated
  #
  # Returns the Compound Annual Growth Rate
    cagr <- round(((finish/start)^(1 /periods) - 1)*100, digits=1)    
  return(cagr)
}

sangam.cumsum <-sangam %>%
  group_by(field) %>%
  mutate(CumSum=cumsum(count))

cagr <- sangam.cumsum %>%
  filter(year==1989 | year==1998)%>%
  group_by(field)%>%
  mutate(cagr=CAGR(start=min(CumSum), finish=max(CumSum), periods=9)) %>%
  select(field, cagr)

cagr<-unique(cagr)

############## Generate comparable social work CAGR, 1989-1998 #################

sw.cagr <- articles.df %>%
  filter(Year<1999 & Year>1988)%>%
  mutate(CumSum=cumsum(N)) %>%
  mutate(field = "socialWork")

sw.cagr <- sw.cagr %>%
  mutate(cagr=CAGR(start=min(CumSum), finish=max(CumSum), periods=9))%>%
  select(field, cagr)

sw.cagr<-unique(sw.cagr)

cagr<-rbind(cagr, sw.cagr)

cagr$field<-as.character(cagr$field)
cagr$field<-reorder(cagr$field, cagr$cagr)

#############PLOT##################

CAGR <- ggplot(cagr, aes(x=field, y=cagr, fill=field, group =1)) + theme_bw() + 
    geom_bar(width=0.8, stat='identity') + theme(axis.text.x = element_text(size=8),
        title = element_text(size = 10), plot.title=element_text(size=10)) + 
    ylab("Compound annual growth rate (%)") + 
    xlab("") + 
    ggtitle("Figure 4. Compound annual growth rates for \nnumber of published articles (1989 - 1998)") + 
    theme(plot.title = element_text(size = rel(1.5))) + scale_y_continuous(breaks =
        c(seq(0, 40,  15)), limits=c(0, 40)) + theme(axis.title.x =
        element_text(vjust=-0.5), axis.title.y = element_text(vjust=.75)) + 
    scale_fill_manual(values=c("#b2abd2", "#b2abd2", "#fdb863", "#b2abd2",
        "#b2abd2", "#b2abd2", "#b2abd2"))+ theme(legend.position="none") + 
    scale_x_discrete(labels=c("Psychology","Economics","Social \nWork","Sociology","Political \nScience", "History", "Anthropology")) + 
    geom_hline(yintercept=seq(0, 40, 15), col="white", lwd=0.6) + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),  
        panel.border = element_blank(), axis.line = element_line(), 
        axis.text.y = element_text(size = 15),
        axis.text.x = element_text(size = 15),
        axis.title.y = element_text(size = 15), 
        title = element_text(size = 13)
        ) + 
    geom_text(aes(label =cagr ), vjust = -.5)

```